Data Descriptor: Corresponding author journal publishing activities 2014-2018

### 2019-10-21 analysis/paper.Rmd

- `journal_publisher_14_18`: Global journal and publisher data 2014-18
- `rp_jn_14_18`: Corresponding author country affiliations per publisher and journal 2014-18

@@

| Variable           | Description                                                                                                                                                                      | Source                                                                                                                                                     |
| :----------------- | :------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- | :--------------------------------------------------------------------------------------------------------------------------------------------------------- |
| `issn_wos`         | ISSN, a standardized journal id.                                                                                                                                                 | KB Web of Science: `wos_b_2019.issues.issn`                                                                                                                |
| `country_code`     | Country of affiliation corresponding author, represented as ISO 3 code                                                                                                           | KB Web of Science: `wos_b_2019.d_items_authors_institutions.inst_countrycode`                                                                              |
| `publication_year` | Year of publication, obtained from KB Web of Science                                                                                                                             | KB Web of Science: `wos_b_2019.items.pubyear`                                                                                                              |
| `articles`         | Number of original articles and reviews published. Whole counting where internationally co-located corresponding authorships were assigned to each contributing country equally. | KB Web of Science: Grouped counts over `wos_b_2019.issues.issn`, `wos_b_2019.d_items_authors_institutions.inst_countrycode` and `wos_b_2019.items.pubyear` |
| `journal_title`    | Most frequently used journal title in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref                                       | Crossref                                                                                                                                                   |
| `publisher`        | Most frequently used publisher name in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref                                      | Crossref                                                                                                                                                   |
| `oa_journal`       | Is the journal publishing all articles open access without delay (full open access)?                                                                                             | ISSN-GOLD_OA List 3.0                                                                                                                              |
| `issn_l`           | Linking ISSN, a journal id that groups different media of the same serial publication, e.g. ISSN for print with electronic ISSN.                                             | CIEPS                                                                                                                                                      |

@@

The following figure shows the country share per publisher in terms of original articles and reviews published in subscription journals using corresponding author affiliations. The flipped x-axis shows the top 10 countries of affiliations in terms of articles published, while the y-axis depicts the percentage per publisher. The figure indicates, which countries were particularly important for a publisher. Furthermore, it suggests large variations across publishers. Take for instance the publisher Royal Society of Chemistry. Around every third article published in a subscription journal was submitted from an author based in China, while the overall share of China-based authors was significantly lower with `r rp_df %>% filter(oa_journal == FALSE) %>% group_by(country_code) %>% summarise(n = sum(articles)) %>% arrange(desc(n)) %>% mutate(prop = n /sum(n)) %>% filter(country_code == "CHN") %>% .$prop * 100` %.

```{r}
top_10_publisher <- rp_df %>%
  filter(oa_journal == FALSE) %>%
  group_by(publisher) %>%
  summarise(n = sum(articles)) %>%
  arrange(desc(n)) %>%
  head(11) %>%
  .$publisher
top_10_countries <- rp_df %>%
  filter(oa_journal == FALSE) %>%
  group_by(country_code) %>%
  summarise(n = sum(articles)) %>%
  arrange(desc(n)) %>%
  head(10) %>%
  .$country_code
plot_df <- rp_df %>%
  filter(oa_journal == FALSE) %>%
  mutate(fct_publisher = ifelse(publisher %in% top_10_publisher, publisher, "Others")) %>%
  mutate(fct_publisher = ifelse(is.na(fct_publisher), "Others", fct_publisher)) %>%
  mutate(fct_countrcode = ifelse(country_code %in% top_10_countries, country_code, "Others")) %>%
  mutate(fct_countrcode = factor(fct_countrcode, levels = c(top_10_countries, "Others"))) %>%
  group_by(fct_publisher, fct_countrcode) %>%
  summarize(n = sum(articles)) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()
#' plot
ggplot(plot_df, aes(forcats::fct_rev(fct_countrcode), prop, fill = prop)) +
  coord_flip() +
  geom_bar(stat = "identity") +
  facet_wrap(~ fct_relevel(fct_publisher, top_10_publisher), ncol = 3) +
  scale_fill_viridis_c(
    "Country share per publisher (in %)",
    labels = scales::percent_format(accuracy = 5L),
    option = "C"
  ) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L))  +
  labs(
    title = "Who did publish in subscription-based journals from major publishers between 2014 - 2018?",
    x = "Country of Affiliation (Top 10)",
    y = "Share of Original Articles and Reviews per Publisher",
    caption = "Data Sources:  Web of Science in-house database Competence Centre for Bibliometrics for German Science.
    Publisher names were disambiguated using Crossref metadata. OA status obtained from the Bielefeld GOLD OA list.
    Funding: Federal Ministry of Education and Research of Germany (BMBF), Project OAUNI (01PU17023A)",
    subtitle = "Corresponding authors' countries of affiliation per publisher."
  ) +
  theme_minimal(base_family = "Roboto", base_size = 16) +
  theme(plot.margin = margin(30, 10, 30, 10)) +
  theme(panel.grid.minor.y = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.y = element_blank()) +
  theme(panel.border = element_blank()) +
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.text = element_text(size =  9)
  ) -> p
#' save as png
ggsave(
  "figure/publisher_share.png",
  p,
  dpi = 450,
  height = 13,
  width = 17
)
```

>>>>

- `journal_publisher_14_18`: Global journal and publisher data 2014-18
- `rp_publisher_14_18.csv`: Corresponding author country affiliations per publisher and journal 2014-18

@@

|Variable           |Description                                                                                                                                 |Source                                                                        |
|:------------------|:-------------------------------------------------------------------------------------------------------------------------------------------|:-----------------------------------------------------------------------------|
|`country_code`     |Country of affiliation corresponding author, represented as ISO 3 code                                                                      |KB Web of Science: `wos_b_2019.d_items_authors_institutions.inst_countrycode` |
|`publication_year` |Year of publication, obtained from KB Web of Science                                                                                        |KB Web of Science: `wos_b_2019.items.pubyear`                                 |
|`publisher`        |Most frequently used publisher name in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref |Crossref                                                                      |
|`oa_journal`       |Is the journal publishing all articles open access without delay (full open access)?                                                        |ISSN GOLD OA List V3                                                          |
|`n_publications`   |Number of original articles and reviews published                                                                                           |Aggregate data                                                                |
|`n_journals`       |Number of distinct journals with corresponding publication output                                                                           |Aggregate data

@@

The following figure shows the country share per publisher in terms of original articles and reviews published in subscription journals using corresponding author affiliations. The flipped x-axis shows the top 10 countries of affiliations in terms of articles published, while the y-axis depicts the percentage per publisher. The figure indicates, which countries were particularly important for a publisher. Furthermore, it suggests large variations across publishers. Take for instance the publisher Royal Society of Chemistry. Around every third article published in a subscription journal was submitted from an author based in China, while the overall share of China-based authors was significantly lower with `r rp_df %>% filter(oa_journal == FALSE) %>% group_by(country_code) %>% summarise(n = sum(n_publications)) %>% arrange(desc(n)) %>% mutate(prop = n /sum(n)) %>% filter(country_code == "CHN") %>% .$prop * 100` %.

```{r}
top_10_publisher <- rp_df %>%
  filter(oa_journal == FALSE) %>%
  group_by(publisher) %>%
  summarise(n = sum(n_publications)) %>%
  arrange(desc(n)) %>%
  head(11) %>%
  .$publisher
top_10_countries <- rp_df %>%
  filter(oa_journal == FALSE) %>%
  group_by(country_code) %>%
  summarise(n = sum(n_publications)) %>%
  arrange(desc(n)) %>%
  head(10) %>%
  .$country_code
plot_df <- rp_df %>%
  filter(oa_journal == FALSE) %>%
  mutate(fct_publisher = ifelse(publisher %in% top_10_publisher, publisher, "Others")) %>%
  mutate(fct_publisher = ifelse(is.na(fct_publisher), "Others", fct_publisher)) %>%
  mutate(fct_countrcode = ifelse(country_code %in% top_10_countries, country_code, "Others")) %>%
  mutate(fct_countrcode = factor(fct_countrcode, levels = c(top_10_countries, "Others"))) %>%
  group_by(fct_publisher, fct_countrcode) %>%
  summarize(n = sum(n_publications)) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()
#' plot
ggplot(plot_df, aes(forcats::fct_rev(fct_countrcode), prop, fill = prop)) +
  coord_flip() +
  geom_bar(stat = "identity") +
  facet_wrap(~ fct_relevel(fct_publisher, top_10_publisher), ncol = 3) +
  scale_fill_viridis_c(
    "Country share per publisher (in %)",
    labels = scales::percent_format(accuracy = 5L),
    option = "C"
  ) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L))  +
  labs(
    title = "Who did publish in subscription-based journals from major publishers between 2014 - 2018?",
    x = "Country of Affiliation (Top 10)",
    y = "Share of Original Articles and Reviews per Publisher",
    caption = "Data Sources:  Web of Science in-house database Competence Centre for Bibliometrics for German Science.
    Publisher names were disambiguated using Crossref metadata. OA status obtained from the Bielefeld GOLD OA list.
    Funding: Federal Ministry of Education and Research of Germany (BMBF), Project OAUNI (01PU17023A)",
    subtitle = "Corresponding authors' countries of affiliation per publisher."
  ) +
  theme_minimal(base_family = "Roboto", base_size = 16) +
  theme(plot.margin = margin(30, 10, 30, 10)) +
  theme(panel.grid.minor.y = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.y = element_blank()) +
  theme(panel.border = element_blank()) +
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.text = element_text(size =  9)
  ) -> p
#' save as png
ggsave(
  "figure/publisher_share.png",
  p,
  dpi = 450,
  height = 13,
  width = 17
)
```
